First Order Logic of Sparse Random Graphs

### 2020-01-21 Article.tex
Seminario

\begin{theorem} 
	The following are satisfied:
	\begin{itemize}
		\item[(1)] Let $r\in \N$ and let $\mathcal{T}$ be a
		$k$-equivalence class for trees with radii at most $r$.
		Then $\mathrm{Pr}[r,\mathcal{T}]$ exists and is an expression
		in $\Lambda$.
		\item[(2)] Let $\overbar{w}\in \N*$ be a list of different fixed 
		vertices, and let $\overbar{v}\in \N*$ be another list of 
		 Let $r_1,\dots r_k\in \N$ and let
		$\mathcal{T}_1,\dots \mathcal{T}_k$ be 
		$k$-equivalence classes for trees such that
		each $\mathcal{T}_i$ has radii bounded by $r_i$. Let
		$\overbar{v}$ denote a list of fixed vertices 
		$v_1,\dots,v_k,\dots, v_l\in \N$, and let 
		$\phi(\overbar{x})$ be an open sentence on $l$ variables.
		Let $G_n^\prime=G_n\setminus E[\overbar{v}]$ 
	\end{itemize}

>>>>

\begin{theorem} 
	The following are satisfied:
	\begin{itemize}
		\item[(1)] Let $r\in \N$ and let $\mathcal{T}$ be a
		$k$-equivalence class for trees with radii at most $r$.
		Then $\mathrm{Pr}[r,\mathcal{T}]$ exists and is an expression
		in $\Lambda$.
		\item[(2)] Let $\overbar{w}\in \N*$ be a list of different fixed 
		vertices, and let $\phi[\overbar{x}]\in FO[\sigma]$ be an open sentence
		(i.e. with no bounded variables) such that $\overbar{x}$ and $\overbar{w}$
		have the same size. Let $=v_1,\dots,v_k \in \N$ be another 
		list of different vertices that are contained in $\overbar{w}$.
		Let $r_1,\dots,r_k \in \N$ and let $\mathcal{T}_1,\dots \mathcal{T}_k$
		be $k$-equivalence classes for trees such that
		each $\mathcal{T}_i$ has radii bounded by $r_i$. Then
		\[
		\Ln \mathrm{Pr}\big( \bigwedge_{i=1}^k 
		T_{n,i}=T_n(v_i,\overbar{w};r_i) \text{ is a tree and } 
		(T_{n,i},v_i)\in \mathcal{T}_i \, | \, \sigma(\overbar{w})
		\big)= \prod_{i=1}^k \mathrm{Pr}[r_i,\mathcal{T}_i]. \]		
	\end{itemize}
	\begin{proof}
		Let $R$ be an upper bound of the radius of $r$ and the 
		$r_i$'s	in the statement. 
		We will prove (1) and (2) together by induction on $R$. \par
		Assume $R=0$. We start by showing that (1) holds. Recall that
		all trees with radius zero are $k$-equivalent. Thus, if
		$\mathcal{T}$ is the unique $k$-equivalence class of trees
		with radius zero
		and $v\in \N$ is a fixed vertex, then
		 \[
		 \mathrm{Pr}[0;\mathcal{T}] = \Ln \mathrm{Pr}\big( 
		 T:=N^{G_n}(v;0) \text{ is a tree, and } (T,v)\in \mathcal{T} \big)
		 =1, 
		 \]
		Indeed, $N^{G_n}(v;0)$ consists of a single vertex for all $n\geq v$,
		and the above equation follows. The expression $1$ belongs to 
		$\Lambda$, so (1) holds. \par
		The case of (2) is analogous. As $R=0$, then $\mathcal{T}_1=\dots=
		\mathcal{T}_k$ are the unique $k$-equivalence class of trees with
		radius zero. Then, given $\sigma, \overbar{w}, v_1,\dots, v_k$ as
		in the statement,
		\[\Ln \mathrm{Pr}\big( \bigwedge_{i=1}^k 
		T_{n,i}=T_n(v_i,\overbar{w};0) \text{ is a tree and } 
		(T_{n,i},v_i)\in \mathcal{T}_i \, | \, \sigma(\overbar{w})
		\big)=1. \]	
		Because of (1), $\mathrm{Pr}[0,\mathcal{T}_i]=1$ for all $i$'s, 
		and (2) holds. \par
		Now let $R>0$ and assume that both (1) and (2) hold for all 
		lesser values of $R$.

	\end{proof}

### 2020-01-21 Presentacion Sparse Graphs/seminario.tex
Seminario

\begin{frame}{Preliminaries}
\begin{center}
	\large{First order language of graphs}
\end{center}
\begin{columns}
	\begin{column}{0.55\textwidth}
		\begin{itemize}
			\item Variables $x_1, \dots, x_n, \dots$
			\item Connectives $\wedge, \vee$, equality symbol $=$,
			and negation symbol $\neg$.
			\item Quantifiers $\forall, \exists$.
			\item A binary relation symbol $R$
		\end{itemize}
	\end{column}
		\begin{column}{0.45\textwidth}
		\begin{itemize}
			\item Vertices.
			\item ``And", ``or", ``equals", ``not".
			\item ``For all", ``there exists".
			\item Edges $x\sim y$. 
		\end{itemize}
	\end{column}
\end{columns}
~\\
\[ \forall x_1,x_2 \, R(x_1,x_2) \implies \exists x_3 \,
 \neg(x_3=x_1)\wedge \neg(x_3=x_1) \wedge R(x_1, x_3)\]
\end{frame}

\begin{frame}{Preliminaries}
	\Large{Binomial model of random graphs $G(n,p)$.} \\
	\large
	~\\
	Start with vertex set $[n]=\{1,\dots, n\}$ and
	each edge is added with probability $p$ independently. 
	\\~\\
	If $G=([n], E)$:
	\[ Pr(G)= p^{|E|}\cdot (1-p)^{\binom{n}{2}- |E|}. \]

>>>>

\begin{frame}{Preliminaries: The first order logic (FO) of graphs}
	Variables $x_1, \dots, x_n, \dots$ $\rightarrow$ Vertices\\
	Boolean connectives $\wedge, \vee, \neg$ and equality symbol $=$.\\
	Quantifiers $\forall, \exists$.\\
	A binary relation symbol $E$.\\~\\
	It can express properties like:
	\begin{itemize}
		\item The existence of a given subgraph
		\item The existence of a covering set of size $k$
	\end{itemize}
	It cannot express connectivity, $k$-colorability, existence of a Hamiltonian
	path \dots
\end{frame}

\begin{frame}{Preliminaries: The binomial model of random graphs $G(n,p)$}

	Start with vertex set $[n]=\{1,\dots, n\}$ and
	each edge is added with probability $p$ independently. 
	\\~\\
	If $G=([n], E)$:
	\[ Pr(G)= p^{|E|}\cdot (1-p)^{\binom{n}{2}- |E|}. \]
